---
title: "Titan FPL Managers Performance Analysis 2022/23 Season"
author: "Ifechi Mafiana"
date: "2023-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Titan FPL Managers Performance Analysis 2022/23 Season
**As a passionate Fantasy Premier League manager, analyzing the Titan FPL managers 
performance for 2022/2023 season was a project that allowed me to showcase my 
technical, analytical, and statistical skills as a data analyst. Excel, R, and 
Power BI are some of the tools used for this analysis.**

### **Business task:** What are the factors that contribute to a team's success in FPL using Titan managers league as case study 

### Setting up my environment

```{r set up environment}
setwd("C:/Users/MrsMavi/Documents/R Programming/Titan")
```

```{r load packages}
library(tidyverse)
library(dplyr)
library(readxl)
library(janitor)
library(psych)
library(lubridate)
library(ggplot2)

```

### Collect Data and Upload data

```{r import all Titan FPL managers datasets}
mavi_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Mavi_Fc")
whales_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Whales_Fc")
premium_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Premium_Fc")
itsnotyou_its_mee <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Itsnotyou_its_Mee")
expatriate_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Expatriate_FC")
thragg_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Thragg_FC")
impeccable_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Impeccable_FC")
regal_reloaded <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Regal_Reloaded")
andra_lina <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Andra_Lina")
spinel_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Spinel_FC")
king_kiadin <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "King_Kiadin")
marbie_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Marbie_FC")
eclipse_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Eclipse_FC")
team_cuevas <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Team_Cuevas")
club_de_phabz <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Club_De_Phabz")
dont_stress_me <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Dont_Stress_Me")
all_guns_blazing <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "All_Guns_Blazing")
layla_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Layla_FC")
ameenify_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Ameenify_FC")
Wakanda_nino_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Wakanda_Nino_FC")
star_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Star_FC")
apollo_xi <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Apollo_XI")
le_quinze <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Le_Quinze")
fC_creamthug <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "FC_Creamthug")
chibest <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Chibest")
mezzy_united <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "Mezzy_United")
vasog_fc <- read_excel("Titan_Fpl_Manager_2022_2023.xlsx", sheet = "VASOG_FC")

```

### Add manager and team columns
```{r add manager and team columns to datasets}
dataset_names <- c("mavi_fc", "whales_fc", "premium_fc", "itsnotyou_its_mee", "expatriate_fc",
                   "thragg_fc", "impeccable_fc", "regal_reloaded", "andra_lina", "spinel_fc",
                   "king_kiadin", "marbie_fc", "eclipse_fc", "team_cuevas", "club_de_phabz",
                   "dont_stress_me", "all_guns_blazing", "layla_fc", "ameenify_fc",
                   "Wakanda_nino_fc", "star_fc", "apollo_xi", "le_quinze", "fC_creamthug",
                   "chibest", "mezzy_united", "vasog_fc")

team_names <- c("Ifechi Mafiana", "Adewale Aderoloye", "Samuel Odunze","El Tunmbs", 
                "Godsgift Ojukwu", "Nelson Omoike", "Godsblessing Ojukwu", "George Ofuani",
                "Providee Baba", "Adewuni Samuel", "Kelly Ikpeze", "Lukman Ikhuoriah", "Akuma Ukpo",
                "Ugochinyereze Akujobi_Emetuche", "Olamide Fabanwo", "Ebere Nwaribe", "Taiwo Ohenhen", 
                "Maryam Ikhuoriah", "Al_Ameen Ayinde","Keny Elnino", "Precious Oriafoh", "Tochukwu Akaegbusi", 
                "franklyn Olerum", "Joseph Harry" , "VeeVee", "Chimezie Obasi", "Victor Asogwa")

# Loop through each dataset and add the manager and team columns
for (i in seq_along(dataset_names)) {
  dataset_name <- dataset_names[i]
  dataset <- get(dataset_name)             # Get the dataset by name
  dataset$Manager <- team_names[i]         # Add the manager column with the corresponding team name
  dataset$Team <- dataset_name             # Add the team column with the dataset name
  assign(dataset_name, dataset)            # Assign the modified dataset back to the original name
}
```

### Add more columns; experience, game week points difference, and location
```{r add experience, game week points difference and location columns}

# Create a vector of start years for each manager
start_years <- c("2018", "2020", "2017", "2012", "2017",
                 "2021", "2019", "2018", "2021", "2017",
                 "2020", "2020", "2018", "2019", "2018",
                 "2019", "2017", "2022", "2020", "2017",
                 "2017", "2018", "2018", "2017", "2017",
                 "2022", "2017")

# Calculate the total experience for each manager
current_year <- as.integer(format(Sys.Date(), "%Y"))
experience <- current_year - as.integer(start_years)

# Iterate over the dataset names and add the "experience" column to each dataset
for (i in seq_along(dataset_names)) {
  dataset_name <- dataset_names[i]
  dataset <- get(dataset_name)
  dataset$Experience <-  as.numeric(experience)[i]
  assign(dataset_name, dataset)
}

# Function to calculate GWP_diff and return the updated dataset
calculate_GWP_diff <- function(dataset) {
  dataset <- dataset %>%
    mutate(GWP_diff = GWP - lag(GWP))
  return(dataset)
}

# Loop through each dataset name
for (name in dataset_names) {
  # Get the dataset object based on the name
  dataset <- get(name)
  
  # Calculate GWP_diff and update the dataset
  dataset <- calculate_GWP_diff(dataset)
  
  # Assign the updated dataset back to the original object
  assign(name, dataset)
}

# Add Location column to the datasets
datasets <- list(
  mavi_fc = "Nigeria",
  whales_fc = "Barbados",
  premium_fc = "Nigeria",
  itsnotyou_its_mee = "Nigeria",
  expatriate_fc = "Nigeria",
  thragg_fc = "Nigeria",
  impeccable_fc = "Nigeria",
  regal_reloaded = "United Kingdom",
  andra_lina = "Nigeria",
  spinel_fc = "Nigeria",
  king_kiadin = "Nigeria",
  marbie_fc = "Nigeria",
  eclipse_fc = "Nigeria",
  team_cuevas = "Nigeria",
  club_de_phabz = "Nigeria",
  dont_stress_me = "Nigeria",
  all_guns_blazing = "Canada",
  layla_fc = "Nigeria",
  ameenify_fc = "Nigeria",
  Wakanda_nino_fc = "Nigeria",
  star_fc = "Nigeria",
  apollo_xi = "Nigeria",
  le_quinze = "Nigeria",
  fC_creamthug = "Nigeria",
  chibest = "United Kingdom",
  mezzy_united = "Nigeria",
  vasog_fc = "Nigeria"
)

# Loop through each dataset name and location
for (name in names(datasets)) {
  dataset <- get(name)
  dataset <- mutate(dataset, Location = datasets[[name]])
  assign(name, dataset)
}

# Loop through each dataset name
for (name in names(datasets)) {
  dataset <- datasets[[name]]
  datasets[[name]] <- dataset
}

# Add Favorite_club column to the datasets
favorite_clubs <- list(
  mavi_fc = "Arsenal",
  whales_fc = "Arsenal",
  premium_fc = "Chelsea",
  itsnotyou_its_mee = "Arsenal",
  expatriate_fc = "Arsenal",
  thragg_fc = "Arsenal",
  impeccable_fc = "Arsenal",
  regal_reloaded = "Arsenal",
  andra_lina = "Liverpool",
  spinel_fc = "Arsenal",
  king_kiadin = "Manchester United",
  marbie_fc = "Arsenal",
  eclipse_fc = "Manchester United",
  team_cuevas = "Arsenal",
  club_de_phabz = "Chelsea",
  dont_stress_me = "Arsenal",
  all_guns_blazing = "Chelsea",
  layla_fc = "Liverpool",
  ameenify_fc = "Chelsea",
  Wakanda_nino_fc = "Liverpool",
  star_fc = "Chelsea",
  apollo_xi = "Nil",
  le_quinze = "Arsenal",
  fC_creamthug = "Chelsea",
  chibest = "Chelsea",
  mezzy_united = "Manchester United",
  vasog_fc = "Arsenal"
)

# Loop through each dataset name and Favorite_club
for (name in names(favorite_clubs)) {
  dataset <- get(name)
  dataset <- mutate(dataset, Favorite_club = favorite_clubs[[name]])
  assign(name, dataset)
}
```

### Merge datasets
```{r merge managers datasets into one}
merged_data <- bind_rows(mavi_fc, whales_fc, premium_fc, itsnotyou_its_mee, expatriate_fc,
                     thragg_fc, impeccable_fc, regal_reloaded, andra_lina, spinel_fc,
                     king_kiadin, marbie_fc, eclipse_fc, team_cuevas, club_de_phabz,
                     dont_stress_me, all_guns_blazing, layla_fc, ameenify_fc,
                     Wakanda_nino_fc, star_fc, apollo_xi, le_quinze, fC_creamthug,
                     chibest, mezzy_united, vasog_fc)
```

### Remove "GW" prefix from the values in the "GW" column
```{r remove GW prefix from column values in GW column}
merged_data$GW <- as.numeric(gsub("GW", "", merged_data$GW))
```

### Remove column "#" 
```{r remove unwanted column}
titan_data <- merged_data %>%
  select(-"#")
```

### Clean the column names
```{r clean column names}
colnames(titan_data) <- tolower(colnames(titan_data))
```

### view merged dataset
```{r view merged dataset}
head(titan_data)
str(titan_data)
summary(titan_data)
```

### Qualitative analysis

### Create a table total points of managers 
```{r table showing total points of managers}
summary_table <- titan_data %>%
  group_by(manager) %>%
  summarize(total_points = sum(gwp) - sum(tc)) %>%
  arrange(desc(total_points))
print(summary_table)
```

### Create a table of the number of managers by country
```{r table showing managers locations}
distinct_managers <- titan_data %>%
  group_by(location) %>%
  summarise(distinct_managers = n_distinct(manager))
print(distinct_managers)
```

### Plot the distribution of distinct managers by location
```{r}
distinct_managers <- titan_data %>%
  group_by(location) %>%
  summarise(distinct_managers = n_distinct(manager))
ggplot(distinct_managers, aes(x = "", y = distinct_managers, fill = location)) +
  geom_bar(stat = "identity") +
  coord_polar("y", start = 0) +
  labs(x = NULL, y = NULL, title = "Titan FPL Managers 2022/23: Distribution of Distinct Managers by Location") +
  theme_void()  
```

### The titan dataset will be saved and imported into Microsoft Power BI to create dashboard and reports

### Save data to a CSV file
```{r save dataset}
write.csv(titan_data, "C:/Users/MrsMavi/Documents/R Programming/Titan/titan_managers.csv", row.names = FALSE)
```




